---
title: "Has Australia's rising material prosperity since 2010 been matched by its social wellbeing, or have the two divereged, particularly since 2022, relative to comparable OECD nations ?"
format: html
editor: visual
Author: Meghna Anand (SID:540360309), Dyuti Shraboni Ghosh (SID:540364743)
---

## Executive Summary

## 1. Introduction

Australia is frequently regarded as one of the world's most prosperous economies. Rising household incomes and wealth suggest Australians are materially better off than they were two decades ago. However, economic prosperity does not necessarily translate into a higher quality of life. This study investigates whether Australia's increasing prosperity has been accompanied by improvements in health, environmental quality, education and subjective wellbeing, and compares Australia's performance with similar OECD nations.

## 2. Data Cleaning and Wrangling

```{r}
#load packages
library(tidyverse)
library(janitor)
library(scales)
library(broom)

#set seed for reproducibility
set.seed(123)

#Load OECD Data
oecd_data <- read_csv(
  "OECD Data.csv",
  show_col_types = FALSE
) |>
  clean_names()

wellbeing <- oecd_data |>
  filter(
    age == "_T",
    sex == "_T",
    education_lev == "_T"
  ) |>
  transmute(
    country_code = ref_area,
    country = reference_area,
    indicator_code = measure,
    indicator = measure_2,
    domain_code = domain,
    domain = domain_2,
    year = as.integer(time_period),
    value = as.numeric(obs_value),
    unit = unit_of_measure,
    status = observation_status
  ) |>
  filter(
    !is.na(year),
    !is.na(value)
  )


# Quick overview
glimpse(wellbeing)

wellbeing |>
  summarise(
    Countries = n_distinct(country),
    Domains = n_distinct(domain),
    Indicators = n_distinct(indicator),
    Years = n_distinct(year)
  )

```

## 3. Exploratory Data Analysis

```{r}
# OECD Quality of Life domains
qol_domains <- c(
  "Health",
  "Knowledge and skills",
  "Environmental quality",
  "Subjective well-being",
  "Safety"
)

wellbeing |>
  filter(domain %in% qol_domains) |>
  distinct(domain, indicator) |>
  arrange(domain, indicator)

wellbeing |>
  distinct(domain) |>
  arrange(domain)

```

**Constructing the quality of life index**

```{r}
# Define the Quality of Life indicators

qol_indicators <- tribble(
  ~domain, ~indicator, ~direction,
  
  "Health",
  "Life expectancy at birth",
  "positive",
  
  "Health",
  "Deaths from suicide, alcohol, drugs",
  "negative",
  
  "Knowledge and skills",
  "Student mathematics skills",
  "positive",
  
  "Environmental quality",
  "Exposed to air pollution",
  "negative",
  
  "Environmental quality",
  "Exposure to extreme temperature",
  "negative",
  
  "Subjective well-being",
  "Life satisfaction",
  "positive",
  
  "Subjective well-being",
  "Negative affect balance",
  "negative"
)

qol_data <- wellbeing |>
  inner_join(
    qol_indicators,
    by = c("domain", "indicator")
  )


qol_scores <- qol_data |>
  group_by(indicator) |>
  mutate(
    score = (value - min(value, na.rm = TRUE)) /
      (max(value, na.rm = TRUE) - min(value, na.rm = TRUE)),
    
    score = if_else(
      direction == "negative",
      1 - score,
      score
    ),
    
    score = score * 100
  ) |>
  ungroup()

domain_scores <- qol_scores |>
  group_by(country, year, domain) |>
  summarise(
    domain_score = mean(score),
    .groups = "drop"
  )


qol_index <- domain_scores |>
  group_by(country, year) |>
  summarise(
    qol_index = mean(domain_score),
    .groups = "drop"
  )
```

**Has Australia's Quality of Life improved?**

```{r}
australia_qol <- qol_index |>
  filter(country == "Australia") |>
  arrange(year)

australia_qol |>
  summarise(
    Start = first(qol_index),
    End = last(qol_index),
    Change = End - Start,
    Percent_Change = 100 * Change / Start,
    Average = mean(qol_index)
  )

trend_model <- lm(qol_index ~ year,
                  data = australia_qol)

tidy(trend_model)
glance(trend_model)

ggplot(australia_qol,
       aes(year, qol_index)) +
  geom_line(linewidth = 1.3,
            colour = "#006BA4") +
  geom_point(size = 2.5,
             colour = "#006BA4") +
  geom_smooth(method = "lm",
              se = FALSE,
              linetype = 2,
              colour = "grey40") +
  labs(
    title = "Australia's Quality of Life has steadily improved",
    subtitle = "Overall Quality of Life Index",
    x = NULL,
    y = "Index (0–100)"
  ) +
  theme_minimal(base_size = 13)
confint(trend_model)
summary(trend_model)
```

**What has driven Australia's Quality of Life?**

```{r}
australia_domains <- domain_scores |>
  filter(country == "Australia")

domain_change <- australia_domains |>
  group_by(domain) |>
  summarise(
    Start = first(domain_score),
    End = last(domain_score),
    Change = End - Start
  ) |>
  arrange(desc(Change))

domain_change

ggplot(domain_change,
       aes(Change,
           reorder(domain, Change),
           fill = Change > 0)) +
  geom_col(show.legend = FALSE) +
  labs(
    title = "Which domains have improved the most?",
    x = "Change in score",
    y = NULL
  ) +
  theme_minimal()

ggplot(australia_domains,
       aes(year,
           domain_score,
           colour = domain)) +
  geom_line(linewidth = 1) +
  facet_wrap(~domain) +
  theme_minimal()
```

```{r}
#| label: plot-australia-indicator-changes
#| echo: false
#| warning: false
#| message: false

indicator_change <- qol_scores |>
  filter(country == "Australia") |>
  group_by(indicator) |>
  arrange(year, .by_group = TRUE) |>
  summarise(
    Start = first(score),
    End = last(score),
    Change = End - Start,
    .groups = "drop"
  ) |>
  arrange(desc(Change))

ggplot(
  indicator_change,
  aes(
    x = Change,
    y = reorder(indicator, Change)
  )
) +
  geom_col() +
  labs(
    title = "Change in Australia's quality-of-life indicators",
    x = "Change in normalised score",
    y = NULL
  ) +
  theme_minimal(base_size = 12)
```

```{r}
#| label: plot-australia-indicator-trends
#| echo: false
#| warning: false
#| message: false

ggplot(
<<<<<<< HEAD
  filter(
    qol_scores,
    country == "Australia"
  ),
  aes(
    x = year,
    y = score,
    group = indicator
  )
) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 1.8) +
  facet_wrap(
    ~indicator,
    scales = "free_x"
  ) +
  labs(
    title = "Australia's quality-of-life indicators over time",
    x = NULL,
    y = "Normalised score"
  ) +
  theme_minimal(base_size = 12)
```

## Australian household income over time

The OECD dataset provides three Australian indicators under the Income and Wealth domain: household disposable income, the top-to-bottom household income ratio, and median net wealth. Household net adjusted disposable income per person was selected as the primary measure of material prosperity because it has a complete annual series from 2010 to 2024 and is suitable for comparison across countries. The other two indicators are retained as supporting context because they contain fewer observations and end in 2020.


```{r}
#| label: prepare-australia-income-data
#| include: false

australia_income <- wellbeing |>
  filter(
    country == "Australia",
    stringr::str_detect(
      indicator,
      stringr::regex(
        "net adjusted disposable income",
        ignore_case = TRUE
      )
    )
  ) |>
  arrange(year) |>
  mutate(
    annual_change = value - lag(value),
    annual_change_pct =
      100 * (value / lag(value) - 1)
  )
```
### Income trend

```{r}
#| label: plot-australia-income
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 5

ggplot(
  australia_income,
  aes(x = year, y = value)
) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_x_continuous(
    breaks = seq(
      min(australia_income$year),
      max(australia_income$year),
      by = 2
    )
  ) +
  scale_y_continuous(
    labels = scales::label_dollar(
      prefix = "$",
      big.mark = ","
    )
  ) +
  labs(
    title = "Australia's household disposable income has increased",
    subtitle = "Net adjusted disposable income per person, 2023 USD PPP",
    x = NULL,
    y = "Income per person"
  ) +
  theme_minimal(base_size = 12)
```

Australia’s household disposable income increased from approximately \$44,625 per person in 2010 to \$50,629 in 2024, an overall increase of 13.5%. However, the trend was not continuous: income peaked at \$52,430 in 2021 before declining in subsequent years.

### Summary of the income trend

```{r}
#| label: summarise-australia-income
#| include: false

australia_income_summary <- australia_income |>
  summarise(
    first_year = first(year),
    latest_year = last(year),
    starting_income = first(value),
    latest_income = last(value),
    total_change = last(value) - first(value),
    percentage_change = 100 * (last(value) / first(value) - 1),
    lowest_income = min(value),
    lowest_year = year[which.min(value)],
    highest_income = max(value),
    highest_year = year[which.max(value)]
  ) |>
  mutate(
    percentage_change = round(percentage_change, 1)
  )

australia_income_summary
```

Australia’s household disposable income increased from \$44,625 per person in 2010 to \$50,629 in 2024, representing an overall increase of 13.5%. However, growth was uneven: income peaked at \$52,430 in 2021, declined in 2022 and 2023, and partially recovered in 2024.

## International income comparison

```{r}
#| label: prepare-international-income-comparison
#| include: false

# Check how many countries have income data in each year
income_year_coverage <- wellbeing |>
  filter(indicator_code == "1_1") |>
  distinct(country_code, year) |>
  count(
    year,
    name = "countries_available"
  ) |>
  arrange(year)

# Use 2022 because it is the latest year with near-complete coverage
comparison_year <- 2022

comparison_year_coverage <- income_year_coverage |>
  filter(year == comparison_year) |>
  pull(countries_available)

maximum_country_coverage <- max(
  income_year_coverage$countries_available
)

comparison_coverage_pct <- round(
  100 * comparison_year_coverage / maximum_country_coverage,
  1
)

# Create the international income dataset for 2022
income_comparison <- wellbeing |>
  filter(
    indicator_code == "1_1",
    year == comparison_year
  ) |>
  distinct(
    country_code,
    country,
    .keep_all = TRUE
  ) |>
  transmute(
    country_code,
    country,
    year,
    income_per_person = value
  ) |>
  arrange(desc(income_per_person)) |>
  mutate(
    income_rank = row_number()
  )
```

The international income comparison uses `r comparison_year`, the most recent year with near-complete country coverage. Household disposable income data were available for `r comparison_year_coverage` countries, compared with a maximum of `r maximum_country_coverage` countries in earlier years. This represents `r comparison_coverage_pct`% of the maximum available coverage and provides a recent comparison while minimising the exclusion of countries due to missing data.

### Australia's international income position

```{r}
#| label: calculate-australia-income-position
#| include: false

australia_income_position <- income_comparison |>
  filter(country_code == "AUS")

australia_income_value <- australia_income_position |>
  pull(income_per_person)

australia_income_rank <- australia_income_position |>
  pull(income_rank)

international_income_median <- median(
  income_comparison$income_per_person,
  na.rm = TRUE
)

difference_from_median_pct <- round(
  100 * (
    australia_income_value /
      international_income_median - 1
  ),
  1
)

australia_income_peers <- income_comparison |>
  mutate(
    distance_from_australia =
      abs(income_per_person - australia_income_value),
    country_group = if_else(
      country_code == "AUS",
      "Australia",
      "Similar-income country"
    )
  ) |>
  slice_min(
    order_by = distance_from_australia,
    n = 9,
    with_ties = FALSE
  ) |>
  arrange(income_per_person)
```

In `r comparison_year`, Australia recorded household disposable income of `r scales::dollar(australia_income_value, accuracy = 1)` per person and ranked `r australia_income_rank` out of `r nrow(income_comparison)` countries. Australia's income was approximately `r difference_from_median_pct`% above the median of the countries included in the comparison.

### Australia’s similar-income peer group

To create a focused economic benchmark, similar-income countries were defined as those whose household disposable income was within 15% of Australia’s 2022 value. This comparison helps assess whether Australia achieves stronger or weaker quality-of-life outcomes than countries with broadly comparable household resources.

```{r}
#| label: create-australia-income-peer-group
#| include: false

peer_band_pct <- 0.15

peer_lower_bound <- australia_income_value *
  (1 - peer_band_pct)

peer_upper_bound <- australia_income_value *
  (1 + peer_band_pct)

australia_income_peers <- income_comparison |>
  filter(
    income_per_person >= peer_lower_bound,
    income_per_person <= peer_upper_bound
  ) |>
  mutate(
    country_group = if_else(
      country_code == "AUS",
      "Australia",
      "Similar-income country"
    )
  ) |>
  arrange(income_per_person)

peer_country_count <- australia_income_peers |>
  filter(country_code != "AUS") |>
  nrow()
```

Using this definition, Australia was compared with `r peer_country_count` countries whose household disposable income fell between `r scales::dollar(peer_lower_bound, accuracy = 1)` and `r scales::dollar(peer_upper_bound, accuracy = 1)` per person.

```{r}
#| label: plot-australia-similar-income-countries
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 6

ggplot(
  australia_income_peers,
  aes(
    x = reorder(country, income_per_person),
    y = income_per_person,
    fill = country_group
  )
) +
  geom_col() +
  geom_text(
    aes(
      label = scales::dollar(
        income_per_person,
        accuracy = 1,
        big.mark = ","
      )
    ),
    hjust = -0.1,
    size = 3.4
  ) +
  coord_flip(
    clip = "off"
  ) +
  scale_y_continuous(
    labels = scales::label_dollar(
      accuracy = 1,
      big.mark = ","
    ),
    expand = expansion(
      mult = c(0, 0.2)
    )
  ) +
  labs(
    title = paste0(
      "Countries within 15% of Australia's household income, ",
      comparison_year
    ),
    subtitle = "Net adjusted disposable income per person, 2023 USD PPP",
    x = NULL,
    y = "Income per person",
    fill = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "none",
    plot.margin = margin(
      t = 5.5,
      r = 40,
      b = 5.5,
      l = 5.5
    )
  )
```

These countries will form the focused peer group for comparing Australia’s quality-of-life performance. However, all countries with complete and comparable data will be retained for the overall international ranking and regression analysis.

## Preliminary economic finding

Australia’s household disposable income increased by 13.5% between 2010 and 2024, indicating that households became materially better off on this measure. However, progress was uneven: income peaked in 2021, declined in 2022 and 2023, and partially recovered in 2024. Internationally, Australia ranked sixth out of 36 countries in 2022 and recorded income 20.6% above the sample median. Thirteen countries fell within 15% of Australia’s income and will provide a focused peer group for evaluating whether Australia achieves comparatively stronger or weaker quality-of-life outcomes despite having similar household resources.
=======
  filter(qol_scores,
         country=="Australia"),
  aes(year,
      score,
      colour=indicator)
)+
  geom_line()+
  facet_wrap(~indicator,
             scales="free_y")

```
>>>>>>> 3d529cc (Add quality of life predictive modelling)

## 4. Relationship Analysis

## 5. Predictive Modelling

```{r}
#identify prosperity indicators
prosperity_candidates <- wellbeing |>
  distinct(domain, indicator) |>
  filter(
    str_detect(
      str_to_lower(paste(domain, indicator)),
      "income|wealth|employment|unemployment|earnings|housing|financial|labour|labor|economic|gdp"
    )
  ) |>
  arrange(domain, indicator)

prosperity_candidates |>
  print(n = Inf)

```

```{r}
# CREATE MODELLING DATASET
prosperity_indicators <- c(
  "Households and NPISHs net adjusted disposable income per capita",
  "Median net wealth",
  "Employment rate",
  "Housing affordability"
)

prosperity_data <- wellbeing |>
  filter(indicator %in% prosperity_indicators) |>
  mutate(
    indicator = case_when(
      indicator == "Households and NPISHs net adjusted disposable income per capita" ~ "income",
      indicator == "Median net wealth" ~ "wealth",
      indicator == "Employment rate" ~ "employment",
      indicator == "Housing affordability" ~ "housing"
    )
  ) |>
  select(country, year, indicator, value) |>
  pivot_wider(
    names_from = indicator,
    values_from = value
  )

ml_data <- qol_index |>
  inner_join(
    prosperity_data,
    by = c("country", "year")
  ) |>
  drop_na()

glimpse(ml_data)

ml_data |>
  summarise(
    Observations = n(),
    Countries = n_distinct(country),
    Years = n_distinct(year)
  )

colSums(is.na(ml_data))

```

```{r}
# ---------------------------------------------------------
# STEP 3: CORRELATION ANALYSIS
# ---------------------------------------------------------

cor_matrix <- ml_data |>
  select(
    qol_index,
    income,
    wealth,
    housing,
    employment
  ) |>
  cor()

round(cor_matrix, 2)

library(corrplot)
corrplot(
  cor_matrix,
  method = "color",
  type = "upper",
  addCoef.col = "black",
  tl.col = "black",
  tl.srt = 45,
  number.cex = 0.7
)

cor.test(ml_data$qol_index, ml_data$income)

cor.test(ml_data$qol_index, ml_data$wealth)

cor.test(ml_data$qol_index, ml_data$housing)

cor.test(ml_data$qol_index, ml_data$employment)

```

```{r}
## MULTIPLE LINEAR REGRESSION

full_model <- lm(
  qol_index ~ income + wealth + housing + employment,
  data = ml_data
)

summary(full_model)

library(broom)

tidy(full_model, conf.int = TRUE)

glance(full_model)

tidy(full_model, conf.int = TRUE) |>
  filter(term != "(Intercept)") |>
  ggplot(aes(
    x = estimate,
    y = reorder(term, estimate)
  )) +
  geom_vline(
    xintercept = 0,
    linetype = "dashed",
    colour = "grey60"
  ) +
  geom_point(size = 3, colour = "#006BA4") +
  geom_errorbarh(
    aes(
      xmin = conf.low,
      xmax = conf.high
    ),
    height = 0.2,
    colour = "#006BA4"
  ) +
  labs(
    title = "Estimated effect of prosperity indicators on Quality of Life",
    x = "Regression coefficient",
    y = NULL
  ) +
  theme_minimal(base_size = 13)

summary(full_model)
tidy(full_model, conf.int = TRUE)
glance(full_model)
```

- Overall model is highly significant (F-test, *p* \< 0.001)

- Adjusted R\^2 = **0.427**, so prosperity explains about **43%** of the variation in QoL.

- Employment is the strongest predictor.

- Income is also significant.

- Wealth is borderline (*p* = 0.058).

- Housing affordability doesn't appear to contribute much once the other variables are included.

```{r}
# ---------------------------------------------------------
# STEPWISE MODEL SELECTION
# ---------------------------------------------------------

library(MASS)

step_model <- stepAIC(
  full_model,
  direction = "both",
  trace = FALSE
)

summary(step_model)
AIC(full_model, step_model)
anova(step_model, full_model)
tidy(step_model, conf.int = TRUE)
```

```{r}
# ---------------------------------------------------------
# STEP 6:CHECKING REGRESSION ASSUMPTIONS
# ---------------------------------------------------------

par(mfrow = c(2, 2))
plot(step_model)
par(mfrow = c(1, 1))

shapiro.test(residuals(step_model))

library(lmtest)

bptest(step_model)

library(car)
vif(step_model)

```

```{r}
training_data <- ml_data |>
  filter(country != "Australia")

australia_data <- ml_data |>
  filter(country == "Australia")

# ---------------------------------------------------------
# FINAL MODEL (TRAINED WITHOUT AUSTRALIA)
# ---------------------------------------------------------

final_model <- lm(
  qol_index ~ income + wealth + housing + employment,
  data = training_data
)

summary(final_model)


library(broom)

tidy(final_model, conf.int = TRUE)

glance(final_model)
```

```{r}
australia_predictions <- australia_data |>
  mutate(
    predicted_qol = predict(
      final_model,
      newdata = australia_data
    ),
    prediction_error = qol_index - predicted_qol
  )

australia_predictions
```

```{r}
australia_summary <- australia_predictions |>
  summarise(
    mean_actual = mean(qol_index),
    mean_predicted = mean(predicted_qol),
    mean_error = mean(prediction_error),
    rmse = sqrt(mean(prediction_error^2))
  )

australia_summary

```

```{r}
library(ggplot2)
library(tidyr)
library(dplyr)
library(scales)

plot_data <- australia_predictions |>
  dplyr::select(year, qol_index, predicted_qol) |>
  tidyr::pivot_longer(
    cols = c(qol_index, predicted_qol),
    names_to = "series",
    values_to = "qol"
  ) |>
  dplyr::mutate(
    series = dplyr::recode(
      series,
      qol_index = "Actual QoL",
      predicted_qol = "Expected QoL from prosperity model"
    )
  )
ggplot(
  plot_data,
  aes(
    x = factor(year),
    y = qol,
    group = series,
    colour = series,
    linetype = series
  )
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 3) +
  scale_y_continuous(
    limits = c(0, 100),
    breaks = seq(0, 100, 20),
    expand = expansion(mult = c(0, 0.03))
  ) +
  scale_linetype_manual(
    values = c(
      "Actual QoL" = "solid",
      "Expected QoL from prosperity model" = "dashed"
    )
  ) +
  labs(
    title = "Australia's actual and expected Quality of Life",
    subtitle = paste0(
      "Expected values were estimated using income, wealth, ",
      "housing affordability and employment data from other OECD countries"
    ),
    x = "Year",
    y = "Quality of Life Index (0–100)",
    colour = NULL,
    linetype = NULL,
    caption = paste0(
      "Only years with complete data for all four prosperity indicators are shown. ",
      "Mean actual QoL = 68.28; mean expected QoL = 73.30."
    )
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold"),
    plot.caption = element_text(hjust = 0)
  )
```

## 6. Limitations

## 7. Conclusion

## 8. References
