---
title: "Has Australia's rising material prosperity since 2010 translated into comparable improvements in social wellbeing?"

author:
  - "Meghna Anand (SID: 540360309)"
  - "Dyuti Shraboni Ghosh (SID: 540364743)"

date: today
date-format: "D MMMM YYYY"

format:
  html:
    toc: true

editor: visual

execute:
  echo: false
  warning: false
  message: false
---

## Executive Summary
Australia is one of the world's most prosperous economies, with high household incomes, substantial wealth and strong employment outcomes. However, rising prosperity does not necessarily translate into improvements in broader measures of social wellbeing. This report investigates whether Australia's increasing material prosperity since 2010 has been accompanied by comparable improvements in quality of life.

A composite Quality-of-Life Index was constructed using seven OECD indicators covering health, knowledge and skills, environmental quality, and subjective wellbeing. To enable consistent historical comparison, a Core Quality-of-Life Index was developed using the four indicators available across every three-year period.

Australia's household disposable income increased by **13.5%** between 2010 and 2024, while the Core Quality-of-Life Index declined over the same period. This suggests that improvements in material prosperity were not matched by comparable improvements in social wellbeing.

To investigate this relationship further, a multiple regression model was estimated using OECD country-period observations. The model found that prosperity indicators and period controls accounted for **44.1%** of the variation in quality of life, while Australia consistently recorded lower quality of life than predicted from its prosperity measures.

Overall, the analysis suggests that although material prosperity contributes to quality of life, economic growth alone has not translated into comparable improvements in Australia's social wellbeing.

## 1. Introduction

Australia has experienced sustained growth in material prosperity over the past decade, reflected by high household incomes, wealth and employment. However, these measures do not fully capture people's quality of life, which also depends on health, environmental quality, education and subjective wellbeing.

This report investigates whether Australia's rising material prosperity since 2010 has translated into comparable improvements in social wellbeing.

The analysis first examines Australia's trends in household disposable income and quality of life over time. A Quality-of-Life Index is then constructed to measure broader wellbeing outcomes. Finally, a multiple regression model is used to examine the relationship between prosperity and quality of life across OECD countries, allowing Australia's wellbeing to be compared with the level expected from its prosperity.

## 2. Data Cleaning and Wrangling

The analysis uses the OECD Current Well-being dataset, which contains internationally comparable measures of prosperity and social wellbeing across OECD countries.

The dataset was filtered to retain observations for the total population only. Variables relating to country, indicator, wellbeing domain, year, observed value, unit of measurement and observation status were retained, while records with missing values were removed. Duplicate observations, indicator definitions and units of measurement were checked to ensure consistent comparisons across countries and years.

Following data cleaning, the analytical dataset contained **8,806 observations** across **47 countries**, **10 wellbeing domains**, **21 indicators** and **17 years**.

To measure social wellbeing, a composite Quality-of-Life Index was constructed using seven OECD indicators. Because indicator availability differed over time, a Core Quality-of-Life Index containing the four indicators consistently available across every three-year period was developed for Australia's historical analysis. Household disposable income was used as the primary measure of material prosperity because it provides a complete annual series from 2010 to 2024.

```{r}
#load packages
library(tidyverse)
library(janitor)
library(scales)
library(broom)

#set seed for reproducibility
set.seed(123)

#Load OECD Data
oecd_data <- read_csv(
  "OECD Data.csv",
  show_col_types = FALSE
) |>
  clean_names()

wellbeing <- oecd_data |>
  filter(
    age == "_T",
    sex == "_T",
    education_lev == "_T"
  ) |>
  transmute(
    country_code = ref_area,
    country = reference_area,
    indicator_code = measure,
    indicator = measure_2,
    domain_code = domain,
    domain = domain_2,
    year = as.integer(time_period),
    value = as.numeric(obs_value),
    unit = unit_of_measure,
    status = observation_status
  ) |>
  filter(
    !is.na(year),
    !is.na(value)
  )


# Quick overview
glimpse(wellbeing)

wellbeing |>
  summarise(
    Countries = n_distinct(country),
    Domains = n_distinct(domain),
    Indicators = n_distinct(indicator),
    Years = n_distinct(year)
  )

```

## 3. Exploratory Data Analysis
### Constructing the Quality-of-Life Index

Quality of life is multidimensional and cannot be represented by a single OECD indicator. A broader Quality-of-Life Index was therefore constructed using seven indicators spanning health, knowledge and skills, environmental quality and subjective well-being. Each indicator was normalised using min–max scaling to produce scores between 0 and 100. Adverse indicators were reverse-scored so that higher values consistently represented better outcomes. Indicators were first averaged within domains, and the domain scores were then given equal weight in the overall index.

Because indicator availability differed across years, the broader index was not suitable for direct historical comparison. Australia’s longitudinal trend was therefore assessed using a Core Quality-of-Life Index containing the four indicators available in every three-year period. This improves comparability over time, although the core index covers only health, environmental quality and subjective well-being.

```{r}
#| label: construct-quality-of-life-index
#| include: false

qol_indicators <- tribble(
  ~domain, ~indicator, ~direction,

  "Health",
  "Life expectancy at birth",
  "positive",

  "Health",
  "Deaths from suicide, alcohol, drugs",
  "negative",

  "Knowledge and skills",
  "Student mathematics skills",
  "positive",

  "Environmental quality",
  "Exposed to air pollution",
  "negative",

  "Environmental quality",
  "Exposure to extreme temperature",
  "negative",

  "Subjective well-being",
  "Life satisfaction",
  "positive",

  "Subjective well-being",
  "Negative affect balance",
  "negative"
)

qol_data <- wellbeing |>
  inner_join(
    qol_indicators,
    by = c("domain", "indicator")
  )

qol_scores <- qol_data |>
  group_by(indicator) |>
  mutate(
    indicator_min = min(value, na.rm = TRUE),
    indicator_max = max(value, na.rm = TRUE),

    score = case_when(
      indicator_max == indicator_min ~ NA_real_,

      direction == "positive" ~
        100 * (
          value - indicator_min
        ) / (
          indicator_max - indicator_min
        ),

      direction == "negative" ~
        100 * (
          indicator_max - value
        ) / (
          indicator_max - indicator_min
        )
    )
  ) |>
  ungroup() |>
  filter(!is.na(score))

domain_scores <- qol_scores |>
  group_by(
    country,
    year,
    domain
  ) |>
  summarise(
    domain_score = mean(
      score,
      na.rm = TRUE
    ),
    indicators_available = n_distinct(indicator),
    .groups = "drop"
  )

qol_index <- domain_scores |>
  group_by(
    country,
    year
  ) |>
  summarise(
    qol_index = mean(
      domain_score,
      na.rm = TRUE
    ),
    domains_available = n_distinct(domain),
    .groups = "drop"
  )
```

```{r}
#| label: audit-australia-qol-coverage
#| include: false

australia_qol_coverage <- qol_scores |>
  filter(country == "Australia") |>
  group_by(year) |>
  summarise(
    domains_available = n_distinct(domain),
    indicators_available = n_distinct(indicator),
    indicators_included = paste(
      sort(unique(indicator)),
      collapse = "; "
    ),
    .groups = "drop"
  )
```

```{r}
#| label: construct-period-quality-of-life-index
#| include: false

qol_period_data <- qol_scores |>
  filter(year <= 2024) |>
  mutate(
    period = case_when(
      between(year, 2010, 2012) ~ "2010–2012",
      between(year, 2013, 2015) ~ "2013–2015",
      between(year, 2016, 2018) ~ "2016–2018",
      between(year, 2019, 2021) ~ "2019–2021",
      between(year, 2022, 2024) ~ "2022–2024",
      TRUE ~ NA_character_
    ),
    period = factor(
      period,
      levels = c(
        "2010–2012",
        "2013–2015",
        "2016–2018",
        "2019–2021",
        "2022–2024"
      )
    )
  ) |>
  filter(!is.na(period))

qol_period_indicator_scores <- qol_period_data |>
  group_by(
    country,
    period,
    domain,
    indicator
  ) |>
  summarise(
    indicator_score = mean(
      score,
      na.rm = TRUE
    ),
    .groups = "drop"
  )

qol_period_domain_scores <- qol_period_indicator_scores |>
  group_by(
    country,
    period,
    domain
  ) |>
  summarise(
    domain_score = mean(
      indicator_score,
      na.rm = TRUE
    ),
    indicators_available = n_distinct(indicator),
    .groups = "drop"
  )

qol_period_index <- qol_period_domain_scores |>
  group_by(
    country,
    period
  ) |>
  summarise(
    qol_index = mean(
      domain_score,
      na.rm = TRUE
    ),
    domains_available = n_distinct(domain),
    .groups = "drop"
  )

australia_qol_period_coverage <-
  qol_period_indicator_scores |>
  filter(country == "Australia") |>
  group_by(period) |>
  summarise(
    domains_available = n_distinct(domain),
    indicators_available = n_distinct(indicator),
    indicators_included = paste(
      sort(unique(indicator)),
      collapse = "; "
    ),
    .groups = "drop"
  )
```

```{r}
#| label: audit-missing-qol-period-components
#| include: false

expected_domains <- unique(qol_indicators$domain)
expected_indicators <- unique(qol_indicators$indicator)

australia_qol_period_coverage_detailed <-
  qol_period_indicator_scores |>
  filter(country == "Australia") |>
  group_by(period) |>
  summarise(
    domains_available = n_distinct(domain),
    indicators_available = n_distinct(indicator),

    missing_domains = paste(
      setdiff(
        expected_domains,
        unique(domain)
      ),
      collapse = "; "
    ),

    missing_indicators = paste(
      setdiff(
        expected_indicators,
        unique(indicator)
      ),
      collapse = "; "
    ),

    .groups = "drop"
  ) |>
  mutate(
    missing_domains = if_else(
      missing_domains == "",
      "None",
      missing_domains
    ),
    missing_indicators = if_else(
      missing_indicators == "",
      "None",
      missing_indicators
    )
  )
```

```{r}
#| label: construct-core-period-qol-index
#| include: false

number_of_periods <- n_distinct(
  qol_period_indicator_scores$period
)

common_period_indicators <-
  qol_period_indicator_scores |>
  filter(country == "Australia") |>
  distinct(period, indicator) |>
  count(
    indicator,
    name = "periods_available"
  ) |>
  filter(
    periods_available == number_of_periods
  ) |>
  pull(indicator)

australia_core_period_domains <-
  qol_period_indicator_scores |>
  filter(
    country == "Australia",
    indicator %in% common_period_indicators
  ) |>
  group_by(
    country,
    period,
    domain
  ) |>
  summarise(
    domain_score = mean(
      indicator_score,
      na.rm = TRUE
    ),
    indicators_available = n_distinct(indicator),
    .groups = "drop"
  )

australia_core_period_index <-
  australia_core_period_domains |>
  group_by(
    country,
    period
  ) |>
  summarise(
    core_qol_index = mean(
      domain_score,
      na.rm = TRUE
    ),
    domains_available = n_distinct(domain),
    indicators_available = sum(indicators_available),
    .groups = "drop"
  )
```

### Trends in Australia’s core quality of life

```{r}
#| label: summarise-australia-core-qol-trend
#| include: false

australia_core_period_summary <-
  australia_core_period_index |>
  arrange(period) |>
  summarise(
    start_period = as.character(first(period)),
    end_period = as.character(last(period)),
    start_index = first(core_qol_index),
    end_index = last(core_qol_index),
    change = end_index - start_index,
    percentage_change =
      100 * change / start_index
  )
```

```{r}
#| label: plot-australia-core-qol-trend
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 5

ggplot(
  australia_core_period_index,
  aes(
    x = period,
    y = core_qol_index,
    group = 1
  )
) +
  geom_line(linewidth = 1) +
  geom_point(size = 2.8) +
  geom_text(
    aes(
      label = scales::number(
        core_qol_index,
        accuracy = 0.1
      )
    ),
    vjust = -0.8,
    size = 3.5
  ) +
  scale_y_continuous(
    limits = c(0, 100),
    breaks = seq(0, 100, by = 20)
  ) +
  labs(
    title = "Australia’s core quality-of-life performance declined modestly",
    subtitle = "Consistent four-indicator index across three-year periods",
    x = NULL,
    y = "Core Quality-of-Life Index (0–100)"
  ) +
  theme_minimal(base_size = 12)
```

Using a consistent set of four indicators available in every period, Australia’s Core Quality-of-Life Index declined from `r scales::number(australia_core_period_summary$start_index, accuracy = 0.1)` in `r australia_core_period_summary$start_period` to `r scales::number(australia_core_period_summary$end_index, accuracy = 0.1)` in `r australia_core_period_summary$end_period`. This represents a decrease of `r scales::number(abs(australia_core_period_summary$change), accuracy = 0.1)` points, or approximately `r scales::number(abs(australia_core_period_summary$percentage_change), accuracy = 0.1)`%. Most of the decline occurred before 2019–2021, after which the index remained broadly stable. This suggests that Australia’s rising material prosperity was not accompanied by an improvement in these consistently measured quality-of-life outcomes.

The Core Quality-of-Life Index covers health, environmental quality and subjective well-being. Knowledge and skills was excluded from the historical trend because student mathematics data were not available in every period.

```{r}
#| label: calculate-core-domain-change
#| include: false

core_domain_change <-
  australia_core_period_domains |>
  arrange(domain, period) |>
  group_by(domain) |>
  summarise(
    start_period = as.character(first(period)),
    end_period = as.character(last(period)),
    start_score = first(domain_score),
    end_score = last(domain_score),
    change = end_score - start_score,
    .groups = "drop"
  ) |>
  arrange(desc(change))
```

### What has driven Australia’s core quality of life?

```{r}
#| label: plot-core-domain-change
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 4.8

ggplot(
  core_domain_change,
  aes(
    x = change,
    y = reorder(domain, change),
    fill = change > 0
  )
) +
  geom_col(
    show.legend = FALSE
  ) +
  geom_vline(
    xintercept = 0,
    linetype = "dashed"
  ) +
  geom_text(
    aes(
      label = scales::number(
        change,
        accuracy = 0.1
      ),
      hjust = if_else(
        change >= 0,
        -0.15,
        1.15
      )
    ),
    size = 3.5
  ) +
  scale_x_continuous(
    expand = expansion(
      mult = c(0.15, 0.15)
    )
  ) +
  labs(
    title = "Change in Australia’s consistently measured domains",
    subtitle = "2010–2012 compared with 2022–2024",
    x = "Change in normalised domain score",
    y = NULL
  ) +
  theme_minimal(base_size = 12)
```

Australia’s consistently measured quality-of-life domains moved in different directions between 2010–2012 and 2022–2024. Environmental quality remained broadly stable, while health declined by approximately 3.9 points and subjective well-being declined by 8.0 points. The overall reduction in the Core Quality-of-Life Index was therefore driven primarily by weaker subjective well-being and, to a lesser extent, health outcomes. Knowledge and skills was excluded because student mathematics data were not available throughout the full study period.

## Australian household income over time

The OECD dataset provides three Australian indicators under the Income and Wealth domain: household disposable income, the top-to-bottom household income ratio, and median net wealth. Household net adjusted disposable income per person was selected as the primary measure of material prosperity because it has a complete annual series from 2010 to 2024 and is suitable for comparison across countries. The other two indicators are retained as supporting context because they contain fewer observations and end in 2020.

```{r}
#| label: prepare-australia-income-data
#| include: false

australia_income <- wellbeing |>
  filter(
    country == "Australia",
    stringr::str_detect(
      indicator,
      stringr::regex(
        "net adjusted disposable income",
        ignore_case = TRUE
      )
    )
  ) |>
  arrange(year) |>
  mutate(
    annual_change = value - lag(value),
    annual_change_pct =
      100 * (value / lag(value) - 1)
  )
```

### Income trend

```{r}
#| label: plot-australia-income
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 5

ggplot(
  australia_income,
  aes(x = year, y = value)
) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_x_continuous(
    breaks = seq(
      min(australia_income$year),
      max(australia_income$year),
      by = 2
    )
  ) +
  scale_y_continuous(
    labels = scales::label_dollar(
      prefix = "$",
      big.mark = ","
    )
  ) +
  labs(
    title = "Australia's household disposable income has increased",
    subtitle = "Net adjusted disposable income per person, 2023 USD PPP",
    x = NULL,
    y = "Income per person"
  ) +
  theme_minimal(base_size = 12)
```

Australia’s household disposable income increased from approximately \$44,625 per person in 2010 to \$50,629 in 2024, an overall increase of 13.5%. However, the trend was not continuous: income peaked at \$52,430 in 2021 before declining in subsequent years.

### Summary of the income trend

```{r}
#| label: summarise-australia-income
#| include: false

australia_income_summary <- australia_income |>
  summarise(
    first_year = first(year),
    latest_year = last(year),
    starting_income = first(value),
    latest_income = last(value),
    total_change = last(value) - first(value),
    percentage_change = 100 * (last(value) / first(value) - 1),
    lowest_income = min(value),
    lowest_year = year[which.min(value)],
    highest_income = max(value),
    highest_year = year[which.max(value)]
  ) |>
  mutate(
    percentage_change = round(percentage_change, 1)
  )

australia_income_summary
```

Australia’s household disposable income increased from \$44,625 per person in 2010 to \$50,629 in 2024, representing an overall increase of 13.5%. However, growth was uneven: income peaked at \$52,430 in 2021, declined in 2022 and 2023, and partially recovered in 2024.

## International income comparison

```{r}
#| label: prepare-international-income-comparison
#| include: false

# Check how many countries have income data in each year
income_year_coverage <- wellbeing |>
  filter(indicator_code == "1_1") |>
  distinct(country_code, year) |>
  count(
    year,
    name = "countries_available"
  ) |>
  arrange(year)

# Use 2022 because it is the latest year with near-complete coverage
comparison_year <- 2022

comparison_year_coverage <- income_year_coverage |>
  filter(year == comparison_year) |>
  pull(countries_available)

maximum_country_coverage <- max(
  income_year_coverage$countries_available
)

comparison_coverage_pct <- round(
  100 * comparison_year_coverage / maximum_country_coverage,
  1
)

# Create the international income dataset for 2022
income_comparison <- wellbeing |>
  filter(
    indicator_code == "1_1",
    year == comparison_year
  ) |>
  distinct(
    country_code,
    country,
    .keep_all = TRUE
  ) |>
  transmute(
    country_code,
    country,
    year,
    income_per_person = value
  ) |>
  arrange(desc(income_per_person)) |>
  mutate(
    income_rank = row_number()
  )
```

The international income comparison uses `r comparison_year`, the most recent year with near-complete country coverage. Household disposable income data were available for `r comparison_year_coverage` countries, compared with a maximum of `r maximum_country_coverage` countries in earlier years. This represents `r comparison_coverage_pct`% of the maximum available coverage and provides a recent comparison while minimising the exclusion of countries due to missing data.

### Australia's international income position

```{r}
#| label: calculate-australia-income-position
#| include: false

australia_income_position <- income_comparison |>
  filter(country_code == "AUS")

australia_income_value <- australia_income_position |>
  pull(income_per_person)

australia_income_rank <- australia_income_position |>
  pull(income_rank)

international_income_median <- median(
  income_comparison$income_per_person,
  na.rm = TRUE
)

difference_from_median_pct <- round(
  100 * (
    australia_income_value /
      international_income_median - 1
  ),
  1
)

australia_income_peers <- income_comparison |>
  mutate(
    distance_from_australia =
      abs(income_per_person - australia_income_value),
    country_group = if_else(
      country_code == "AUS",
      "Australia",
      "Similar-income country"
    )
  ) |>
  slice_min(
    order_by = distance_from_australia,
    n = 9,
    with_ties = FALSE
  ) |>
  arrange(income_per_person)
```

In `r comparison_year`, Australia recorded household disposable income of `r scales::dollar(australia_income_value, accuracy = 1)` per person and ranked `r australia_income_rank` out of `r nrow(income_comparison)` countries. Australia's income was approximately `r difference_from_median_pct`% above the median of the countries included in the comparison.

### Australia’s similar-income peer group

To create a focused economic benchmark, similar-income countries were defined as those whose household disposable income was within 15% of Australia’s 2022 value. This comparison helps assess whether Australia achieves stronger or weaker quality-of-life outcomes than countries with broadly comparable household resources.

```{r}
#| label: create-australia-income-peer-group
#| include: false

peer_band_pct <- 0.15

peer_lower_bound <- australia_income_value *
  (1 - peer_band_pct)

peer_upper_bound <- australia_income_value *
  (1 + peer_band_pct)

australia_income_peers <- income_comparison |>
  filter(
    income_per_person >= peer_lower_bound,
    income_per_person <= peer_upper_bound
  ) |>
  mutate(
    country_group = if_else(
      country_code == "AUS",
      "Australia",
      "Similar-income country"
    )
  ) |>
  arrange(income_per_person)

peer_country_count <- australia_income_peers |>
  filter(country_code != "AUS") |>
  nrow()
```

Using this definition, Australia was compared with `r peer_country_count` countries whose household disposable income fell between `r scales::dollar(peer_lower_bound, accuracy = 1)` and `r scales::dollar(peer_upper_bound, accuracy = 1)` per person.

```{r}
#| label: plot-australia-similar-income-countries
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 6

ggplot(
  australia_income_peers,
  aes(
    x = reorder(country, income_per_person),
    y = income_per_person,
    fill = country_group
  )
) +
  geom_col() +
  geom_text(
    aes(
      label = scales::dollar(
        income_per_person,
        accuracy = 1,
        big.mark = ","
      )
    ),
    hjust = -0.1,
    size = 3.4
  ) +
  coord_flip(
    clip = "off"
  ) +
  scale_y_continuous(
    labels = scales::label_dollar(
      accuracy = 1,
      big.mark = ","
    ),
    expand = expansion(
      mult = c(0, 0.2)
    )
  ) +
  labs(
    title = paste0(
      "Countries within 15% of Australia's household income, ",
      comparison_year
    ),
    subtitle = "Net adjusted disposable income per person, 2023 USD PPP",
    x = NULL,
    y = "Income per person",
    fill = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "none",
    plot.margin = margin(
      t = 5.5,
      r = 40,
      b = 5.5,
      l = 5.5
    )
  )
```

These countries will form the focused peer group for comparing Australia’s quality-of-life performance. However, all countries with complete and comparable data will be retained for the overall international ranking and regression analysis.

## Preliminary economic finding

Australia’s household disposable income increased by 13.5% between 2010 and 2024, indicating that households became materially better off on this measure. However, progress was uneven: income peaked in 2021, declined in 2022 and 2023, and partially recovered in 2024. Internationally, Australia ranked sixth out of 36 countries in 2022 and recorded income 20.6% above the sample median. Thirteen countries fell within 15% of Australia’s income and will provide a focused peer group for evaluating whether Australia achieves comparatively stronger or weaker quality-of-life outcomes despite having similar household resources.

## 4. Material prosperity and quality of life

This section examines whether material prosperity is associated with quality of life across OECD countries. To preserve comparability, the analysis uses the same four quality-of-life indicators and the same three-year periods for every included country-period observation.

```{r}
#| label: construct-international-core-qol-index
#| include: false

# Identify country-periods containing all four core QoL indicators
complete_core_qol_cases <-
  qol_period_indicator_scores |>
  filter(
    indicator %in% common_period_indicators
  ) |>
  group_by(
    country,
    period
  ) |>
  summarise(
    indicators_available = n_distinct(indicator),
    domains_available = n_distinct(domain),
    .groups = "drop"
  ) |>
  filter(
    indicators_available ==
      length(common_period_indicators),
    domains_available == 3
  )

# Retain only complete country-period observations
international_core_indicator_scores <-
  qol_period_indicator_scores |>
  filter(
    indicator %in% common_period_indicators
  ) |>
  inner_join(
    complete_core_qol_cases |>
      select(
        country,
        period
      ),
    by = c(
      "country",
      "period"
    )
  )

# Average indicators within their domains
international_core_domain_scores <-
  international_core_indicator_scores |>
  group_by(
    country,
    period,
    domain
  ) |>
  summarise(
    domain_score = mean(
      indicator_score,
      na.rm = TRUE
    ),
    .groups = "drop"
  )

# Give each represented domain equal weight
international_core_qol <-
  international_core_domain_scores |>
  group_by(
    country,
    period
  ) |>
  summarise(
    core_qol_index = mean(
      domain_score,
      na.rm = TRUE
    ),
    domains_available = n_distinct(domain),
    .groups = "drop"
  )
```

```{r}
#| label: construct-period-prosperity-data
#| include: false

prosperity_indicators <- tribble(
  ~indicator, ~prosperity_variable,

  "Households and NPISHs net adjusted disposable income per capita",
  "income",

  "Median net wealth",
  "wealth",

  "Housing affordability",
  "housing",

  "Employment rate",
  "employment"
)

# Assign prosperity observations to the same three-year periods
prosperity_period_data <- wellbeing |>
  inner_join(
    prosperity_indicators,
    by = "indicator"
  ) |>
  filter(year <= 2024) |>
  mutate(
    period = case_when(
      between(year, 2010, 2012) ~ "2010–2012",
      between(year, 2013, 2015) ~ "2013–2015",
      between(year, 2016, 2018) ~ "2016–2018",
      between(year, 2019, 2021) ~ "2019–2021",
      between(year, 2022, 2024) ~ "2022–2024",
      TRUE ~ NA_character_
    ),
    period = factor(
      period,
      levels = c(
        "2010–2012",
        "2013–2015",
        "2016–2018",
        "2019–2021",
        "2022–2024"
      )
    )
  ) |>
  filter(!is.na(period)) |>
  group_by(
    country,
    period,
    prosperity_variable
  ) |>
  summarise(
    prosperity_value = mean(
      value,
      na.rm = TRUE
    ),
    .groups = "drop"
  ) |>
  pivot_wider(
    names_from = prosperity_variable,
    values_from = prosperity_value
  )

# Join the comparable QoL and prosperity measures
international_model_data <- international_core_qol |>
  left_join(
    prosperity_period_data,
    by = c(
      "country",
      "period"
    )
  )

# Retain country-periods with all four prosperity indicators
international_model_complete <-
  international_model_data |>
  filter(
    if_all(
      c(
        income,
        wealth,
        housing,
        employment
      ),
      ~ !is.na(.x)
    )
  )
```

### Correlation analysis

A correlation matrix provides an initial description of the relationships between the Core Quality-of-Life Index and the four prosperity indicators. These are unadjusted correlations and do not account for relationships among the prosperity measures.

```{r}
#| label: plot-prosperity-qol-correlations
#| echo: false
#| warning: false
#| message: false
#| fig-width: 7
#| fig-height: 6

correlation_data <- international_model_complete |>
  select(
    core_qol_index,
    income,
    wealth,
    housing,
    employment
  )

names(correlation_data) <- c(
  "Core QoL",
  "Income",
  "Wealth",
  "Housing",
  "Employment"
)

correlation_matrix <- cor(
  correlation_data,
  use = "complete.obs"
)

corrplot::corrplot(
  correlation_matrix,
  method = "color",
  type = "upper",
  addCoef.col = "black",
  tl.col = "black",
  tl.srt = 45,
  number.cex = 0.8,
  diag = TRUE
)
```

The correlation matrix describes each relationship separately. A multivariable model is therefore used next to estimate the adjusted association of each prosperity measure while holding the other included measures and period constant.

### Multivariable association model

A multiple linear regression model was estimated using non-Australian country-period observations. Australia was excluded from model estimation so that its observed quality-of-life performance can later be compared with the level predicted from other OECD countries.

Income and wealth were log-transformed because their distributions were strongly right-skewed. All four prosperity measures were then standardised, allowing their estimated associations to be compared on a common scale. Period controls were included, and country-clustered standard errors were used because countries could contribute observations from multiple periods.

```{r}
#| label: fit-and-prepare-association-model
#| include: false

# Estimate the model using countries other than Australia
model_training_raw <- international_model_complete |>
  filter(country != "Australia") |>
  mutate(
    log_income = log(income),
    log_wealth = log(wealth)
  )

# Calculate training-sample means and standard deviations
prosperity_scaling <- model_training_raw |>
  summarise(
    income_mean = mean(log_income),
    income_sd = sd(log_income),

    wealth_mean = mean(log_wealth),
    wealth_sd = sd(log_wealth),

    housing_mean = mean(housing),
    housing_sd = sd(housing),

    employment_mean = mean(employment),
    employment_sd = sd(employment)
  )

# Standardise the four prosperity measures
model_training_data <- model_training_raw |>
  mutate(
    income_z =
      (log_income - prosperity_scaling$income_mean) /
      prosperity_scaling$income_sd,

    wealth_z =
      (log_wealth - prosperity_scaling$wealth_mean) /
      prosperity_scaling$wealth_sd,

    housing_z =
      (housing - prosperity_scaling$housing_mean) /
      prosperity_scaling$housing_sd,

    employment_z =
      (employment - prosperity_scaling$employment_mean) /
      prosperity_scaling$employment_sd
  )

# Fit the model with three-year-period controls
association_model <- lm(
  core_qol_index ~
    income_z +
    wealth_z +
    housing_z +
    employment_z +
    period,
  data = model_training_data
)

# Calculate country-clustered standard errors
clustered_vcov <- sandwich::vcovCL(
  association_model,
  cluster = ~country,
  type = "HC1"
)

clustered_test <- lmtest::coeftest(
  association_model,
  vcov. = clustered_vcov
)

clustered_ci <- lmtest::coefci(
  association_model,
  vcov. = clustered_vcov
)

association_results <- tibble(
  term = rownames(clustered_test),
  estimate = clustered_test[, 1],
  std_error = clustered_test[, 2],
  statistic = clustered_test[, 3],
  p_value = clustered_test[, 4],
  conf_low = clustered_ci[, 1],
  conf_high = clustered_ci[, 2]
)

model_fit <- broom::glance(
  association_model
)

# Prepare prosperity coefficients for the graph
association_plot_data <- association_results |>
  filter(
    term %in% c(
      "income_z",
      "wealth_z",
      "housing_z",
      "employment_z"
    )
  ) |>
  mutate(
    predictor = recode(
      term,
      income_z = "Household income",
      wealth_z = "Median wealth",
      housing_z = "Housing affordability",
      employment_z = "Employment rate"
    )
  )

income_result <- association_results |>
  filter(term == "income_z")

wealth_result <- association_results |>
  filter(term == "wealth_z")

housing_result <- association_results |>
  filter(term == "housing_z")

employment_result <- association_results |>
  filter(term == "employment_z")
```



```{r}
#| label: predict-australia-core-qol
#| include: false

# Prepare Australia's prosperity measures using the same
# transformations as the international training data
australia_prediction_data <-
  international_model_complete |>
  filter(country == "Australia") |>
  mutate(
    log_income = log(income),
    log_wealth = log(wealth),

    income_z =
      (log_income - prosperity_scaling$income_mean) /
      prosperity_scaling$income_sd,

    wealth_z =
      (log_wealth - prosperity_scaling$wealth_mean) /
      prosperity_scaling$wealth_sd,

    housing_z =
      (housing - prosperity_scaling$housing_mean) /
      prosperity_scaling$housing_sd,

    employment_z =
      (employment - prosperity_scaling$employment_mean) /
      prosperity_scaling$employment_sd
  )

# Predict Australia's Core QoL using the model trained
# on other OECD countries
australia_prediction_data$predicted_core_qol <-
  predict(
    association_model,
    newdata = australia_prediction_data
  )

# Calculate the difference between actual and predicted QoL
australia_prediction_data <-
  australia_prediction_data |>
  mutate(
    prediction_gap =
      core_qol_index - predicted_core_qol
  )

# Summarise Australia's performance across available periods
australia_prediction_summary <-
  australia_prediction_data |>
  summarise(
    periods_available = n(),
    mean_actual = mean(core_qol_index),
    mean_predicted = mean(predicted_core_qol),
    mean_gap = mean(prediction_gap),
    rmse = sqrt(
      mean(prediction_gap^2)
    )
  )
```


### Is Australia performing as well as its prosperity would suggest?

```{r}
#| label: prepare-australia-prediction-plot
#| include: false

australia_prediction_plot_data <-
  australia_prediction_data |>
  select(
    period,
    core_qol_index,
    predicted_core_qol
  ) |>
  pivot_longer(
    cols = c(
      core_qol_index,
      predicted_core_qol
    ),
    names_to = "series",
    values_to = "qol"
  ) |>
  mutate(
    series = recode(
      series,
      core_qol_index = "Actual Core QoL",
      predicted_core_qol =
        "Expected Core QoL from prosperity"
    )
  )
```

```{r}
#| label: plot-australia-actual-expected-qol
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 5

ggplot(
  australia_prediction_plot_data,
  aes(
    x = period,
    y = qol,
    group = series,
    linetype = series,
    shape = series
  )
) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_text(
    aes(
      label = scales::number(
        qol,
        accuracy = 0.1
      ),
      vjust = if_else(
        series == "Expected Core QoL from prosperity",
        -1,
        1.7
      )
    ),
    size = 3.4,
    show.legend = FALSE
  ) +
  scale_y_continuous(
    limits = c(79, 88),
    breaks = seq(80, 88, by = 2),
    expand = expansion(mult = c(0.03, 0.08))
  ) +
  scale_linetype_manual(
    values = c(
      "Actual Core QoL" = "solid",
      "Expected Core QoL from prosperity" = "dashed"
    )
  ) +
  labs(
    title = "Australia’s quality of life remained below its predicted level",
    subtitle = "Expected values estimated from comparable OECD countries; zoomed axis",
    x = NULL,
    y = "Core Quality-of-Life Index",
    linetype = NULL,
    shape = NULL,
    caption = "The vertical axis is restricted to 79–88 to make the difference between the two series visible."
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    plot.caption = element_text(
      hjust = 0,
      size = 9
    )
  )
```
Across the four periods with complete prosperity data, Australia’s average actual Core Quality-of-Life Index was `r scales::number(australia_prediction_summary$mean_actual, accuracy = 0.1)`, compared with a model-predicted average of `r scales::number(australia_prediction_summary$mean_predicted, accuracy = 0.1)`. Australia therefore performed approximately `r scales::number(abs(australia_prediction_summary$mean_gap), accuracy = 0.1)` points below the level associated with its income, wealth, employment and housing measures.

The gap increased from approximately `r scales::number(abs(first(australia_prediction_data$prediction_gap)), accuracy = 0.1)` points in 2010–2012 to `r scales::number(abs(last(australia_prediction_data$prediction_gap)), accuracy = 0.1)` points in 2019–2021. This suggests that Australia’s material prosperity was increasingly accompanied by weaker quality-of-life outcomes than the model would predict from comparable OECD countries. This is an association rather than evidence that prosperity directly caused the difference.

```{r}
#| label: plot-association-coefficients
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 5

ggplot(
  association_plot_data,
  aes(
    x = estimate,
    y = reorder(predictor, estimate)
  )
) +
  geom_vline(
    xintercept = 0,
    linetype = "dashed",
    colour = "grey50"
  ) +
  geom_errorbar(
    aes(
      xmin = conf_low,
      xmax = conf_high
    ),
    orientation = "y",
    width = 0.15,
    linewidth = 0.8
  ) +
  geom_point(size = 3) +
  labs(
    title = "Employment had the strongest adjusted association with quality of life",
    subtitle = "Standardised coefficients with 95% country-clustered confidence intervals",
    x = "Change in Core Quality-of-Life Index",
    y = NULL
  ) +
  theme_minimal(base_size = 12)
```

The model was estimated using `r model_fit$nobs` non-Australian country-period observations. Together, the prosperity measures and period controls accounted for approximately `r scales::percent(model_fit$adj.r.squared, accuracy = 0.1)` of the observed variation in the Core Quality-of-Life Index.

A one-standard-deviation increase in the employment rate was associated with a `r scales::number(employment_result$estimate, accuracy = 0.1)`-point higher Core Quality-of-Life Index. Its 95% confidence interval ranged from `r scales::number(employment_result$conf_low, accuracy = 0.1)` to `r scales::number(employment_result$conf_high, accuracy = 0.1)`.

Income and wealth also had positive estimated associations, but their confidence intervals included zero. The available data therefore do not provide sufficiently precise evidence of independent associations after employment, housing affordability and period were taken into account.

The housing-affordability measure had a negative adjusted association of `r scales::number(housing_result$estimate, accuracy = 0.1)` points. This counterintuitive result should be interpreted cautiously. It represents a conditional association after accounting for the other included prosperity measures and does not imply that lower housing affordability improves quality of life.


## 6. Limitations

- QoL index depends on selected indicators
- Missing OECD observations reduce sample size
- Regression measures association not causation
- Other social factors weren't included

## 7. Conclusion

This analysis demonstrates that Australia's material prosperity has continued to improve over the past decade, particularly in terms of household disposable income. However, improvements in quality of life have been considerably more modest and inconsistent. Across OECD countries, prosperity is associated with higher quality of life, explaining approximately 43% of the observed variation. Nevertheless, Australia's quality of life remains below the level predicted by its prosperity indicators. These findings suggest that economic growth alone is insufficient to guarantee improvements in social wellbeing. Policies aimed at improving health, environmental quality and subjective wellbeing may therefore be just as important as increasing national prosperity when seeking to improve Australians' quality of life.

Overall, the analysis provides evidence that Australia's material prosperity has grown faster than its social wellbeing. While Australia remains one of the wealthiest OECD countries, the predictive model indicates that its quality of life is consistently lower than expected for a nation with its level of prosperity. This suggests that future improvements in national wellbeing are likely to depend not only on economic growth, but also on policies that strengthen health, environmental quality, education and subjective wellbeing.

## 8. References

Organisation for Economic Co-operation and Development (2026) *Current well-being*. OECD Data Explorer. Available at:<https://data-explorer.oecd.org/vis?df%5Bds%5D=DisseminateFinalDMZ&df%5Bid%5D=DSD_HSL%40DF_HSL_CWB&df%5Bag%5D=OECD.WISE.WDP> (Accessed: 21 July 2026).

Organisation for Economic Co-operation and Development (2026)*OECD well-being database definitions*. OECD. Available at:<https://www.oecd.org/content/dam/oecd/en/topics/policy-sub-issues/measuring-well-being-and-progress/oecd-well-being-database-definitions.pdf>(Accessed: 21 July 2026).
